On exponential sums arising from the classical doubling method

### 2025-03-31 main.tex
Correcting tori but not actions
Minor correction

Let $\algebraicGroup{A}_n \subset \algebraicGroup{\GL}_n$ be the diagonal torus. We denote for an element $g \in \algebraicGroup{\GL}_n$, $g^{\ast} = w_n \transpose{g}^{-1} w_n$, where $w_n$ is the long Weyl element (see \Cref{subsec:isometry-groups}). Using $\algebraicGroup{A}_n$ we define standard (mostly split) tori in our groups as follows.
\begin{center}
\begin{tabular}{|c|c|}
	\hline 
		$\algebraicGroup{G}$ & $\algebraicGroup{T}_0$ \tabularnewline
		\hline 
		\hline 
		$\algebraicGroup{\UnitaryGroup}_{2n}$ & $\left\{ \begin{pmatrix}
			a\\
			& \involution{\left(a^{\ast}\right)}
		\end{pmatrix}  \mid a \in \algebraicGroup{A}_n \right\}$ \tabularnewline
		\hline 
		$\algebraicGroup{\UnitaryGroup}_{2n+1}$ & $\left\{ \begin{pmatrix}
	a \\
	& t\\
	& & \involution{\left(a^{\ast}\right)}
\end{pmatrix}  \mid a \in \algebraicGroup{A}_n,\,\, t \in \algebraicGroup{U}_1 \right\}$ \tabularnewline
\hline 
	$\algebraicGroup{\SO}_{2n+1}$ & $\left\{ \begin{pmatrix}
		a\\
		& 1\\
		& & a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_n \right\}$ \tabularnewline
	\hline 
	$\algebraicGroup{\Sp}_{2n}$ or $\algebraicGroup{\SO}^{+}_{2n}$ & $\left\{ \begin{pmatrix}
		a\\
		& a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_n \right\}$ \tabularnewline
	\hline 
	$\algebraicGroup{\SO}_{2n}^{-}$ & $\left\{ \begin{pmatrix}
		a\\
		& h\\
		& & a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_{n-1}\,\, h \in \algebraicGroup{\SO}_2^{-} \right\}$ \tabularnewline
	\hline		
	\end{tabular}
\end{center}
Here, $\algebraicGroup{T}_{0}$ is split for all rows except for the last row. We also write down the standard tori for the relevant corresponding similitude groups.
\begin{center}
	\begin{tabular}{|c|c|}
		\hline 
		$\algebraicGroup{G}$ & $\algebraicGroup{T}_0$ \tabularnewline
		\hline 
		\hline 
		$\algebraicGroup{\GSp}_{2n}$ or $\algebraicGroup{\GSO}^{+}_{2n}$ & $\left\{\left(\begin{pmatrix}
			\lambda a\\
			& a^{\ast}
		\end{pmatrix}, \lambda\right) \mid a \in \algebraicGroup{A}_n,\,\, \lambda \in \multiplcativeScheme \right\}$ \tabularnewline
		\hline 
		$\algebraicGroup{\GSO}_{2n}^{-}$ & $\left\{ \left(\begin{pmatrix}
			\algebraicGroup{\aFieldNorm}_{\finiteFieldExtension{2} \slash \finiteField}\left(h\right) a\\
			& h\\
			& & a^{\ast}
		\end{pmatrix}, \algebraicGroup{\aFieldNorm}_{\finiteFieldExtension{2} \slash \finiteField}\left(h\right)\right) \mid a \in \algebraicGroup{A}_{n-1}\,\, h \in \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme} \right\}$ \tabularnewline
		\hline		
	\end{tabular}
\end{center}
Here we realize $\restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme}$ as $$\restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme} = \left\{ \left(\begin{pmatrix}
	a & b\\
	db & a
\end{pmatrix}, \lambda\right) \mid a^2 - db^2 = \lambda \in \multiplcativeScheme \right\},$$
where $d$ is as in \Cref{subsec:isometry-groups}.

@@

\begin{center}
	\begin{tabular}{|c|c|c|c|c|c|} \hline
		$\algebraicGroup{G}$ & $\algebraicGroup{T} \subset \algebraicGroup{G}$ & $\Frobenius$ action on $\CharacterLattice{\algebraicGroup{T}}$ & $\algebraicGroup{G}^{\ast}$ & $\algebraicGroup{T}^{\ast} \subset \algebraicGroup{G}^{\ast}$ \tabularnewline \hline \hline
		$\algebraicGroup{\UnitaryGroup}_n$ & $\multiplcativeScheme^{n}$ & $\left(-1\right)^{\times n}$ & $\algebraicGroup{\UnitaryGroup}_n$ &  $\multiplcativeScheme^{n}$ \tabularnewline \hline		 
		$\algebraicGroup{\SO}_{2n+1}$ & $\multiplcativeScheme^n$ & 1 &$\algebraicGroup{\Sp}_{2n}$ &  $\multiplcativeScheme^n$ \tabularnewline \hline
		$\algebraicGroup{\Sp}_{2n}$ & $\multiplcativeScheme^n$ & 1 &$\algebraicGroup{\SO}_{2n+1}$ & $\multiplcativeScheme^n$ \tabularnewline \hline		 
		$\algebraicGroup{\SO}^{+}_{2n}$ & $\multiplcativeScheme^n$ & 1 &$\algebraicGroup{\SO}^{+}_{2n}$ &  $\multiplcativeScheme^n$ \tabularnewline \hline
		$\algebraicGroup{\SO}^{-}_{2n}$ & $\multiplcativeScheme^{n-1} \times \algebraicGroup{\UnitaryGroup}_1$ & $\left(1\right)^{\times {\left(n-1\right)}} \times \left(-1\right)$ &$\algebraicGroup{\SO}^{-}_{2n}$ &  $\multiplcativeScheme^{n-1} \times \algebraicGroup{\UnitaryGroup}_1$ \tabularnewline \hline
		$\algebraicGroup{\GSp}_{2n}$ & $\multiplcativeScheme \times \multiplcativeScheme^{n}$ & 1 &$\algebraicGroup{GSpin}_{2n+1}$ &  $\multiplcativeScheme \times \multiplcativeScheme^{n}$ \tabularnewline \hline
		$\algebraicGroup{\GSO}^{+}_{2n}$ & $\multiplcativeScheme \times \multiplcativeScheme^{n}$ & 1 & $\algebraicGroup{GSpin}_{2n}^{+}$ &  $\multiplcativeScheme \times \multiplcativeScheme^{n}$ \tabularnewline \hline
		$\algebraicGroup{\GSO}^{-}_{2n}$ & $\multiplcativeScheme^{n-1} \times \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme}$ & $\left(1\right)^{\times {\left(n-1\right)}} \times \sigma^{\ast}$ & $\algebraicGroup{GSpin}_{2n}^{-}$ &  $\multiplcativeScheme^{n-1} \times \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme}$ \tabularnewline \hline
	\end{tabular}
\end{center}
\calvin{correct the tori and how Fr acts on cocharacters and the Weyl group}
For all the pairs $\algebraicGroup{T} \subset \algebraicGroup{G}$ in this list, the Frobenius root $\Frobenius$ of $\algebraicGroup{G}$ acts trivially on the Weyl group $W\left(\algebraicGroup{T}\right)$. In the last row, $\sigma \colon \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme} \to \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme}$ is the automorphism $a + b\sqrt{d} \mapsto a - b\sqrt{d}$ where $d \in \multiplicativegroup{\finiteField}$ is a non-square.

@@

We recall\[
	\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_n} = \Pi_{a \in \mathbb{Z}/n\mathbb{Z}} \left(\algebraicGroup{\GL}_n\right)_a
\]
with Frobenius root action $\Frobenius(g_0,g_1, \hdots, g_{n-1}) = (\Frobenius(g_{n-1}, \Frobenius(g_0), \Frobenius(g_1), \hdots, \Frobenius(g_{n-2}))$.

\begin{remark}
Torus,Theta pairs in the restriction are the same as torus theta pairs before the restriction.
\end{remark}

>>>>

Let $\algebraicGroup{A}_n \subset \algebraicGroup{\GL}_n$ be the diagonal torus. We denote for an element $g \in \algebraicGroup{\GL}_n$, $g^{\ast} = w_n \transpose{g}^{-1} w_n$, where $w_n$ is the long Weyl element (see \Cref{subsec:isometry-groups}). Using $\algebraicGroup{A}_n$ we define standard (mostly split) tori in our groups as follows.
\begin{center}
\begin{tabular}{|c|c|}
	\hline 
		$\algebraicGroup{G}$ & $\algebraicGroup{T}_0$ \tabularnewline
		\hline 
		\hline 
		$\algebraicGroup{\UnitaryGroup}_{2n}$ & $\left\{ \begin{pmatrix}
			a\\
			& \involution{\left(a^{\ast}\right)}
		\end{pmatrix}  \mid a \in \restrictionOfScalars{\quadraticExtension}{\finiteField}{\algebraicGroup{A}_n} \right\}$ \tabularnewline
		\hline 
		$\algebraicGroup{\UnitaryGroup}_{2n+1}$ & $\left\{ \begin{pmatrix}
	a \\
	& t\\
	& & \involution{\left(a^{\ast}\right)}
\end{pmatrix}  \mid a \in \restrictionOfScalars{\quadraticExtension}{\finiteField}{\algebraicGroup{A}_n},\,\, t \in \algebraicGroup{U}_1 \right\}$ \tabularnewline
\hline 
	$\algebraicGroup{\SO}_{2n+1}$ & $\left\{ \begin{pmatrix}
		a\\
		& 1\\
		& & a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_n \right\}$ \tabularnewline
	\hline 
	$\algebraicGroup{\Sp}_{2n}$ or $\algebraicGroup{\SO}^{+}_{2n}$ & $\left\{ \begin{pmatrix}
		a\\
		& a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_n \right\}$ \tabularnewline
	\hline 
	$\algebraicGroup{\SO}_{2n}^{-}$ & $\left\{ \begin{pmatrix}
		a\\
		& h\\
		& & a^{\ast}
	\end{pmatrix} \mid a \in \algebraicGroup{A}_{n-1}\,\, h \in \algebraicGroup{\SO}_2^{-} \right\}$ \tabularnewline
	\hline		
	\end{tabular}
\end{center}
Here, $\algebraicGroup{T}_{0}$ is split for all rows except for the last row. We also write down the standard tori for the relevant corresponding similitude groups.
\begin{center}
	\begin{tabular}{|c|c|}
		\hline 
		$\algebraicGroup{G}$ & $\algebraicGroup{T}_0$ \tabularnewline
		\hline 
		\hline 
		$\algebraicGroup{\GSp}_{2n}$ or $\algebraicGroup{\GSO}^{+}_{2n}$ & $\left\{\left(\begin{pmatrix}
			\lambda a\\
			& a^{\ast}
		\end{pmatrix}, \lambda\right) \mid a \in \algebraicGroup{A}_n,\,\, \lambda \in \multiplcativeScheme \right\}$ \tabularnewline
		\hline 
		$\algebraicGroup{\GSO}_{2n}^{-}$ & $\left\{ \left(\begin{pmatrix}
			\algebraicGroup{\aFieldNorm}_{\finiteFieldExtension{2} \slash \finiteField}\left(h\right) a\\
			& h\\
			& & a^{\ast}
		\end{pmatrix}, \algebraicGroup{\aFieldNorm}_{\finiteFieldExtension{2} \slash \finiteField}\left(h\right)\right) \mid a \in \algebraicGroup{A}_{n-1}\,\, h \in \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme} \right\}$ \tabularnewline
		\hline		
	\end{tabular}
\end{center}
Here we realize $\restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme}$ as $$\restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme} = \left\{ \left(\begin{pmatrix}
	a & b\\
	db & a
\end{pmatrix}, \lambda\right) \mid a^2 - db^2 = \lambda \in \multiplcativeScheme \right\},$$
where $d$ is as in \Cref{subsec:isometry-groups}.

@@

\begin{center}
	\begin{tabular}{|c|c|c|c|c|c|} \hline
		$\algebraicGroup{G}$ & $\algebraicGroup{T} \subset \algebraicGroup{G}$ & $\Frobenius$ action on $\CharacterLattice{\algebraicGroup{T}}$ & $\algebraicGroup{G}^{\ast}$ & $\algebraicGroup{T}^{\ast} \subset \algebraicGroup{G}^{\ast}$ \tabularnewline \hline \hline
		$\algebraicGroup{\UnitaryGroup}_{2n}$ & $\restrictionOfScalars{\quadraticExtension}{\finiteField}{\multiplcativeScheme^{n}}$ & $\left(-1\right)^{\times n}$ & $\algebraicGroup{\UnitaryGroup}_{2n}$ &  $\restrictionOfScalars{\quadraticExtension}{\finiteField}{\multiplcativeScheme^{n}}$ \tabularnewline \hline
		$\algebraicGroup{\UnitaryGroup}_{2n+1}$ & $\restrictionOfScalars{\quadraticExtension}{\finiteField}{\multiplcativeScheme^{n}} \times \algebraicGroup{\UnitaryGroup}_1$ & $\left(-1\right)^{\times n}$ & $\algebraicGroup{\UnitaryGroup}_{2n+1}$ &  $\restrictionOfScalars{\quadraticExtension}{\finiteField}{\multiplcativeScheme^{n} \times \algebraicGroup{\UnitaryGroup}}_1$ \tabularnewline \hline		 				 
		$\algebraicGroup{\SO}_{2n+1}$ & $\multiplcativeScheme^n$ & 1 &$\algebraicGroup{\Sp}_{2n}$ &  $\multiplcativeScheme^n$ \tabularnewline \hline
		$\algebraicGroup{\Sp}_{2n}$ & $\multiplcativeScheme^n$ & 1 &$\algebraicGroup{\SO}_{2n+1}$ & $\multiplcativeScheme^n$ \tabularnewline \hline		 
		$\algebraicGroup{\SO}^{+}_{2n}$ & $\multiplcativeScheme^n$ & 1 &$\algebraicGroup{\SO}^{+}_{2n}$ &  $\multiplcativeScheme^n$ \tabularnewline \hline
		$\algebraicGroup{\SO}^{-}_{2n}$ & $\multiplcativeScheme^{n-1} \times \algebraicGroup{\UnitaryGroup}_1$ & $\left(1\right)^{\times {\left(n-1\right)}} \times \left(-1\right)$ &$\algebraicGroup{\SO}^{-}_{2n}$ &  $\multiplcativeScheme^{n-1} \times \algebraicGroup{\UnitaryGroup}_1$ \tabularnewline \hline
		$\algebraicGroup{\GSp}_{2n}$ & $\multiplcativeScheme \times \multiplcativeScheme^{n}$ & 1 &$\algebraicGroup{GSpin}_{2n+1}$ &  $\multiplcativeScheme \times \multiplcativeScheme^{n}$ \tabularnewline \hline
		$\algebraicGroup{\GSO}^{+}_{2n}$ & $\multiplcativeScheme \times \multiplcativeScheme^{n}$ & 1 & $\algebraicGroup{GSpin}_{2n}^{+}$ &  $\multiplcativeScheme \times \multiplcativeScheme^{n}$ \tabularnewline \hline
		$\algebraicGroup{\GSO}^{-}_{2n}$ & $\multiplcativeScheme^{n-1} \times \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme}$ & $\left(1\right)^{\times {\left(n-1\right)}} \times \sigma^{\ast}$ & $\algebraicGroup{GSpin}_{2n}^{-}$ &  $\multiplcativeScheme^{n-1} \times \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme}$ \tabularnewline \hline
	\end{tabular}
\end{center}
\calvin{correct the tori and how Fr acts on cocharacters and the Weyl group}
For all the pairs $\algebraicGroup{T} \subset \algebraicGroup{G}$ in this list, the Frobenius root $\Frobenius$ of $\algebraicGroup{G}$ acts trivially on the Weyl group $W\left(\algebraicGroup{T}\right)$. In the last row, $\sigma \colon \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme} \to \restrictionOfScalars{\finiteFieldExtension{2}}{\finiteField}{\multiplcativeScheme}$ is the automorphism $a + b\sqrt{d} \mapsto a - b\sqrt{d}$ where $d \in \multiplicativegroup{\finiteField}$ is a non-square.

@@

We recall\[
	\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_n} = \prod_{a \in \mathbb{Z}/m\mathbb{Z}} \left(\algebraicGroup{\GL}_n\right)_a
\]
with Frobenius root action $$\Frobenius(g_0,g_1, \hdots, g_{m-1}) = (\Frobenius(g_{m-1}), \Frobenius(g_0), \Frobenius(g_1), \hdots, \Frobenius(g_{m-2})).$$

\begin{remark}
Torus character pairs for $\restrictionOfScalars{\finiteFieldExtension{m}}{\finiteField}{\algebraicGroup{\GL}_n}$ are the same as torus character pairs for $\algebraicGroup{\GL}_n$ as a group over $\finiteFieldExtension{m}$.
\end{remark}
